---
title: "Code: Mediación ejemplos"
subtitle: "clase 16"
author: "dacarras"
date: "Noviembre 17, 2022"
output: github_document
---


```{r , echo=TRUE, include = FALSE}
# -----------------------------------------------
# general setup
# -----------------------------------------------

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'png')
options(knitr.kable.NA = '')
options(scipen = 9999)

rm(list = ls()) # remove all previous objects

# fonts
Sys.setenv(LANG="en_US.UTF-8")

# -----------------------------------------------
# get times
# -----------------------------------------------

start_time <- Sys.time()

# -----------------------------------------------
# local path
# -----------------------------------------------

# get local path from relative path
local_path <- function(x){
paste0(tools::file_path_as_absolute(x),'/')  
}



```

# Folder locations

```{r, echo=TRUE}

# -------------------------------------------------------------------
# folder locations
# -------------------------------------------------------------------

# -----------------------------------------------
# get local path from relative path
# -----------------------------------------------

local_path <- function(x){
paste0(tools::file_path_as_absolute(x),'/')  
}

# -----------------------------------------------
# dacarras folders
# -----------------------------------------------

data_folder   <- local_path(getwd())
output_folder <- local_path(getwd())

# -----------------------------------------------
# check folders
# -----------------------------------------------

list.files(data_folder)
list.files(output_folder)

# -----------------------------------------------
# load main library
# -----------------------------------------------

library(dplyr)

```


# Read files

```{r, echo=TRUE}

# -------------------------------------------------------------------
# open file
# -------------------------------------------------------------------

data_aut <- readRDS('data_aut.rds')

# -----------------------------------------------
# check data structure
# -----------------------------------------------

dplyr::glimpse(data_aut)


```

# Population Model

## Naive Model

```{r, echo=TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_ch09 <- data_aut %>%
             dplyr::filter(cty %in% c('CHL')) %>%
             dplyr::filter(year == 2009) %>%
             dplyr::select(aut, ses, civ, id_j) %>%
             na.omit()

# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
aut ~ c*ses # y <- x
aut ~ b*civ # y <- m

civ ~ a*ses # m <- x

# effects of interest
xm  := a
my  := b
xy  := c
tot := c + a*b
ind := a*b
dir := c
'

# -----------------------------------------------
# fit model
# -----------------------------------------------

m01 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_ch09)

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m01, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m01) %>%
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)



```

## Clustered Errors

```{r, echo=TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_ch09 <- data_aut %>%
             dplyr::filter(cty %in% c('CHL')) %>%
             dplyr::filter(year == 2009) %>%
             dplyr::select(aut, ses, civ, id_j) %>%
             na.omit()


# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
aut ~ c*ses # y <- x
aut ~ b*civ # y <- m

civ ~ a*ses # m <- x

# effects of interest
xm  := a
my  := b
xy  := c
tot := c + a*b
ind := a*b
dir := c
'

# -----------------------------------------------
# fit model
# -----------------------------------------------

m02 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_ch09,
       cluster = 'id_j') 

# Mote: last line requests clustered errors

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m02, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m02) %>%
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```


## Bootstrapped errors

```{r, echo=TRUE}


# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_ch09 <- data_aut %>%
             dplyr::filter(cty %in% c('CHL')) %>%
             dplyr::filter(year == 2009) %>%
             dplyr::select(aut, ses, civ, id_j) %>%
             na.omit()


# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
aut ~ c*ses # y <- x
aut ~ b*civ # y <- m

civ ~ a*ses # m <- x

# effects of interest
xm  := a
my  := b
xy  := c
tot := c + a*b
ind := a*b
dir := c
'

# -----------------------------------------------
# fit model
# -----------------------------------------------

m03 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_ch09,
       bootstrap = 5000,
       cluster = 'id_j')

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m03, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m03,
boot.ci.type = "bca.simple") %>% 
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```

## Graded Response model

- Can't account for clustered errors simultaneously.

```{r, echo=TRUE}


# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_ch09 <- data_aut %>%
             dplyr::filter(cty %in% c('CHL')) %>%
             dplyr::filter(year == 2009) %>%
             dplyr::select(au1:au9, ses, civ, id_j) %>%
             na.omit()


# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
lau =~ 1*au1
lau =~ 1*au2
lau =~ 1*au3
lau =~ 1*au4
lau =~ 1*au5
lau =~ 1*au6
lau =~ 1*au7
lau =~ 1*au8
lau =~ 1*au9

lau~~var*lau

lau ~ c*ses # y <- x
lau ~ b*civ # y <- m

civ ~ a*ses # m <- x

# effects of interest
xm  := a
my  := b
xy  := c
tot := c + a*b
ind := a*b
dir := c
'

# -----------------------------------------------
# ordinal indicators
# -----------------------------------------------

aut_items <- data_ch09 %>%
             dplyr::select(au1:au9) %>%
             names()

# -----------------------------------------------
# fit model
# -----------------------------------------------

m04 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       estimator = 'WLSMV',
       data = data_ch09,
       ordered= aut_items,
       bootstrap = 5000
       )

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m04, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m04,
boot.ci.type = "bca.simple") %>% 
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```


## Graded Response model (2pl)

- Can't account for clustered errors simultaneously.

```{r, echo=TRUE}


# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_ch09 <- data_aut %>%
             dplyr::filter(cty %in% c('CHL')) %>%
             dplyr::filter(year == 2009) %>%
             dplyr::select(au1:au9, ses, civ, id_j) %>%
             na.omit()


# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
lau =~ NA*au1
lau =~ l2*au2
lau =~ l3*au3
lau =~ l4*au4
lau =~ l5*au5
lau =~ l6*au6
lau =~ l7*au7
lau =~ l8*au8
lau =~ l9*au9

lau~~1*lau

lau ~ c*ses # y <- x
lau ~ b*civ # y <- m

civ ~ a*ses # m <- x

# effects of interest
xm  := a
my  := b
xy  := c
tot := c + a*b
ind := a*b
dir := c
'

# -----------------------------------------------
# ordinal indicators
# -----------------------------------------------

aut_items <- data_ch09 %>%
             dplyr::select(au1:au9) %>%
             names()

# -----------------------------------------------
# fit model
# -----------------------------------------------

m04 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       estimator = 'WLSMV',
       data = data_ch09,
       ordered= aut_items,
       bootstrap = 5000
       )

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m04, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m04,
boot.ci.type = "bca.simple") %>% 
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```

# Multigroup Model

## Free parameters

```{r, echo=TRUE}


# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------


data_ch09_mg <- data_aut %>%
                dplyr::filter(cty %in% c('CHL')) %>%
                dplyr::filter(year == 2009) %>%
                dplyr::select(aut, ses, civ, inx, id_j) %>%
                na.omit()

# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
aut ~ c(c1,c2)*ses # y <- x
aut ~ c(b1,b2)*civ # y <- m

civ ~ c(a1,a2)*ses # m <- x

# free intercepts
aut ~ c(i1, i2)*1
civ ~ c(i3, i4)*1


# indirect effects g1
tot1 := c1 + a1*b1
ind1 := a1*b1
dir1 := c1

## indirect effects g2
tot2 := c2 + a2*b2
ind2 := a2*b2
dir2 := c2
'


# -----------------------------------------------
# fit model
# -----------------------------------------------

m05 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_ch09_mg,
       group = 'inx',
       cluster = 'id_j')

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m05, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m05)%>% 
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)

lavaan::parameterEstimates(m05)%>% 
dplyr::filter(label %in% c('a1','a2')) %>%
knitr::kable(., digits = 2)

lavaan::parameterEstimates(m05)%>% 
dplyr::filter(label %in% c('b1','b2')) %>%
knitr::kable(., digits = 2)

lavaan::parameterEstimates(m05)%>% 
dplyr::filter(label %in% c('tot1','tot2')) %>%
knitr::kable(., digits = 2)


```

## Constrained direct and second stage parameters

```{r, echo=TRUE}


# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------


data_ch09_mg <- data_aut %>%
                dplyr::filter(cty %in% c('CHL')) %>%
                dplyr::filter(year == 2009) %>%
                dplyr::select(aut, ses, civ, inx, id_j) %>%
                na.omit()

# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
aut ~ c(c1,c1)*ses # y <- x
aut ~ c(b1,b1)*civ # y <- m

civ ~ c(a1,a2)*ses # m <- x

# free intercepts
aut ~ c(i1, i2)*1
civ ~ c(i3, i4)*1


'


# -----------------------------------------------
# fit model
# -----------------------------------------------

m05c <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_ch09_mg,
       group = 'inx',
       cluster = 'id_j')

# -----------------------------------------------
# likelihood ratio test
# -----------------------------------------------

lavaan::lavTestLRT(m05, m05c, method = 'satorra.bentler.2001')

# -----------------------------------------------
# conclusion
# -----------------------------------------------

# Note: first stage moderation. 
#       Family's Political interested accelerates
#       the effect of socioeconomic level on civic knowledge.

```

# Multilevel Mediation

## Multilevel latent covariate

- 1-1-1 model (unconflated MLM)
- Preacher, K. J., Zyphur, M. J., & Zhang, Z. (2010). A general multilevel SEM framework for assessing multilevel mediation. Psychological Methods, 15(3), 209–233. https://doi.org/10.1037/a0020141
- [download link](https://www.dropbox.com/s/gxa0hj0o61fu7tg/Preacher%2C%20Zyphur%2C%20Zhang_2010.pdf?dl=1)

```{r, echo=TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_ch09 <- data_aut %>%
             dplyr::filter(cty %in% c('CHL')) %>%
             dplyr::filter(year == 2009) %>%
             dplyr::select(aut, ses, civ, id_j) %>%
             na.omit()


# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
level: 1
aut ~ c1*ses # y <- x
aut ~ b1*civ # y <- m
civ ~ a1*ses # m <- x

level: 2
aut ~ c2*ses # y <- x
aut ~ b2*civ # y <- m
civ ~ a2*ses # m <- x

## L1 effects
xm1  := a1
my1  := b1
xy1  := c1
tot1 := c1 + a1*b1
ind1 := a1*b1
dir1 := c1

## L2 effects
xm2  := a2
my2  := b2
xy2  := c2
tot2 := c2 + a2*b2
ind2 := a2*b2
dir2 := c2
'


# -----------------------------------------------
# fit model
# -----------------------------------------------

m04 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_ch09,
       cluster = 'id_j')

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m04, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m04)%>% 
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```






## Multilevel (1-2-1-1) Open Classroom Discussion

```{r, echo=TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_ch16 <- data_aut %>%
             dplyr::filter(year == 2016) %>%
             dplyr::select(aut, ses, civ, opd, cln, id_j) %>%
             na.omit()


# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
level: 1
aut ~ c11*ses
aut ~ b11*civ
aut ~ e11*opd

civ ~ a11*ses
civ ~ d11*opd

opd ~ f11*ses


level: 2
aut ~ c21*ses
aut ~ b21*civ
aut ~ e21*opd

civ ~ a21*ses
civ ~ d21*opd

opd ~ f21*ses


## indirect effects 2
ind1 := f21*e21
ind2 := f21*d21*b21
ind3 := a21*b21
dir2 := c21
tot2 := c21 + ind1 + ind2 + ind3

'


# -----------------------------------------------
# fit model
# -----------------------------------------------

m04 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_ch16,
       cluster = 'id_j')

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m04, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m04)%>% 
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```
