---
title: "Code: Mediación (OTL) ejemplos"
subtitle: "clase 16"
author: "dacarras"
date: 'Noviembre 17, 2022'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width = 10, fig.asp = 1.15)
```


# Update psi2301 packages

```{r, echo=TRUE}

#------------------------------------------------------------------------------
# update psi2301
#------------------------------------------------------------------------------

# -----------------------------------------------
# get github
# -----------------------------------------------

# credentials::set_github_pat()

# remotes::install_github("dacarras/psi2301")

```

# Prepare data

```{r, echo=TRUE}

#------------------------------------------------------------------------------
# prepare data
#------------------------------------------------------------------------------

# -----------------------------------------------
# load dplyr
# -----------------------------------------------

library(dplyr)


ctry_name_table <- read.table(text = "
IDCNTRY  ctry_name
12  'Algeria'
32  'Argentina'
36  'Australia'
40  'Austria'
48  'Bahrain'
51  'Armenia'
70  'Bosnia and Herzegovina'
72  'Botswana'
76  'Brazil'
84  'Belize'
100 'Bulgaria'
124 'Canada'
152 'Chile'
158 'Chinese Taipei'
170 'Colombia'
196 'Cyprus'
200 'Czech Republic'
203 'Czech Republic'
208 'Denmark'
214 'Dominican Republic'
222 'El Salvador'
233 'Estonia'
246 'Finland'
250 'France'
268 'Georgia'
275 'Palestinian National Authority'
276 'Germany'
288 'Ghana'
300 'Greece'
320 'Guatemala'
344 'Hong Kong, SAR'
348 'Hungary'
352 'Iceland'
360 'Indonesia'
364 'Iran, Islamic Republic of'
372 'Ireland'
376 'Israel'
380 'Italy'
392 'Japan'
398 'Kazakhstan'
400 'Jordan'
410 'Korea, Republic of'
414 'Kuwait'
422 'Lebanon'
428 'Latvia'
438 'Liechtenstein'
440 'Lithuania'
442 'Luxembourg'
458 'Malaysia'
470 'Malta'
484 'Mexico'
496 'Mongolia'
498 'Moldova'
504 'Morocco'
512 'Oman'
528 'Netherlands'
554 'New Zealand'
578 'Norway'
600 'Paraguay'
608 'Philippines'
616 'Poland'
620 'Portugal'
634 'Qatar'
642 'Romania'
643 'Russian Federation'
682 'Saudi Arabia'
702 'Singapore'
703 'Slovak Republic'
705 'Slovenia'
710 'South Africa'
724 'Spain'
752 'Sweden'
756 'Switzerland'
760 'Syria, Arab Republic of'
764 'Thailand'
780 'Trinidad And Tobago'
788 'Tunisia'
792 'Turkey'
804 'Ukraine'
807 'Macedonia'
818 'Egypt'
826 'England'
840 'United States'
887 'Yemen'
891 'Serbia'
926 'England'
927 'Scotland'
956 'Belgium (Flemish)'
957 'Belgium (French)'
3724  'Spain (Basque Country)'
3752  'Sweden (Grade 3)'
4710  'South Africa (Grade 4)'
6162  'Poland (Second-Cycle Programs)'
6431  'Russian Federation (Moscow)'
7241  'Spain (Catalonia)'
7841  'United Arab Emirates (Dubai)'
9132  'Canada (Ontario)'
9133  'Canada (Quebec)'
9134  'Canada (Alberta)'
9135  'Canada (British Columbia)'
9136  'Canada (Novia Scotia)'
9352  'Iceland (Grade 5)'
9578  'Norway (Grade 5)'
9702  'Singapore National'
11800 'United States (Indiana)'
12500 'United States (Massachusetts)'
12700 'United States (Minnesota)'
48401 'Mexico (Generales/Tecnicas/Privadas)'
48402 'Mexico (Telesecundarias)'
48499 'Mexico (Talis-Nacional)'
57891 'Norway (ALU)'
57892 'Norway (ALU +)'
57893 'Norway (PPU)'
57894 'Norway (MASTERS)'
",
head = TRUE)


# -----------------------------------------------
# load data
# -----------------------------------------------

iccs_09 <- psi2301::iccs_2009

# -----------------------------------------------
# variable centering
# -----------------------------------------------

iccs_otl <- iccs_09 %>%
# add country names
dplyr::left_join(., ctry_name_table, by = 'IDCNTRY') %>%
# father tertiary education
mutate(edt_m = case_when(
    MISCED == 0 ~ 0, # not complete ISCED 1
    MISCED == 1 ~ 0, # ISCED 1 primary
    MISCED == 2 ~ 0, # ISCED 2 lower secondary
    MISCED == 3 ~ 0, # ISCED 3 upper secondary
    MISCED == 4 ~ 0, # ISCED 4 post secondary, non tertiary
    MISCED == 5 ~ 1  # ISCED 5A or 6 Tertiary Education
    )) %>%
# highest parent tertiary education
mutate(edt_f = case_when(
    FISCED == 0 ~ 0, # not complete ISCED 1
    FISCED == 1 ~ 0, # ISCED 1 primary
    FISCED == 2 ~ 0, # ISCED 2 lower secondary
    FISCED == 3 ~ 0, # ISCED 3 upper secondary
    FISCED == 4 ~ 0, # ISCED 4 post secondary, non tertiary
    FISCED == 5 ~ 1  # ISCED 5A or 6 Tertiary Education
    )) %>%
mutate(edt_max = pmax(edt_f, edt_m)) %>%
## tertiary education among parents
mutate(edt = edt_max)        %>%  # mean score
mutate(edt_c = r4sda::c_mean(edt, id_j)) %>%  # means by group
mutate(edt_g = r4sda::c_mean(edt, id_k)) %>%  # grand mean             
mutate(edt_w = edt - edt_c   )    %>%  # centering within cluster
mutate(edt_m = edt - edt_g   )    %>%  # centering to the grand mean
mutate(edt_b = edt_c - edt_g )    %>%  # centered cluster means
## ses
mutate(ses = NISB) %>%
mutate(ses = r4sda::z_score(ses))        %>%  # mean score
mutate(ses_c = r4sda::c_mean(ses, id_j)) %>%  # means by group
mutate(ses_g = r4sda::c_mean(ses, id_k)) %>%  # grand mean             
mutate(ses_w = ses - ses_c   )    %>%  # centering within cluster
mutate(ses_m = ses - ses_g   )    %>%  # centering to the grand mean
mutate(ses_b = ses_c - ses_g )    %>%  # centered cluster means
## classroom discussion
mutate(opd = OPDISC)                     %>%  # mean score
mutate(opd = r4sda::z_score(opd))        %>%  # mean score
mutate(opd_c = r4sda::c_mean(opd, id_j)) %>%  # means by group
mutate(opd_g = r4sda::c_mean(opd, id_k)) %>%  # grand mean
mutate(opd_w = opd - opd_c   )    %>%  # centering within cluster
mutate(opd_m = opd - opd_g   )    %>%  # centering to the grand mean
mutate(opd_b = opd_c - opd_g )    %>%  # centered cluster means
## civ
mutate(civ = PV1CIV) %>%
mutate(civ = (civ-500)/100)              %>%  # mean score
mutate(civ_c = r4sda::c_mean(civ, id_j)) %>%  # means by group
mutate(civ_g = r4sda::c_mean(civ, id_k)) %>%  # grand mean                     
mutate(civ_w = civ - civ_c  )  %>%  # centering within cluster
mutate(civ_m = civ - civ_g  )  %>%  # centering to the grand mean
mutate(civ_b = civ_c - civ_g)  %>%  # centered cluster means
# select variables
dplyr::select(ctry_name, id_j, edt:edt_b, ses:ses_b, opd:opd_b, civ:civ_b) %>%
dplyr::glimpse()


```

# Population Model


## Naive Model

- New Zealand

```{r, echo=TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_med <- iccs_otl %>%
            dplyr::filter(ctry_name %in% c('New Zealand')) %>%
            dplyr::select(civ, ses, opd_b, id_j) %>%
            na.omit()

# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
civ ~ c*ses   # y <- x
civ ~ b*opd_b # y <- m

opd_b ~ a*ses # m <- x

# effects of interest
xm  := a
my  := b
xy  := c
tot := c + a*b
ind := a*b
dir := c
pme := ind/tot
'

# -----------------------------------------------
# fit model
# -----------------------------------------------

m01 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_med)

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m01, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m01) %>%
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)

```

## Clustered Errors

```{r, echo=TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_med <- iccs_otl %>%
            dplyr::filter(ctry_name %in% c('New Zealand')) %>%
            dplyr::select(civ, ses, opd_b, id_j) %>%
            na.omit()

# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
civ ~ c*ses   # y <- x
civ ~ b*opd_b # y <- m

opd_b ~ a*ses # m <- x

# effects of interest
xm  := a
my  := b
xy  := c
tot := c + a*b
ind := a*b
dir := c
pme := ind/tot
'

# -----------------------------------------------
# fit model
# -----------------------------------------------

m02 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_med,
       cluster = 'id_j') 

# Mote: last line requests clustered errors

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m02, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m02) %>%
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```

## Bootstrapped errors

```{r, echo=TRUE}


# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------


# -----------------------------------------------
# prep data
# -----------------------------------------------

data_med <- iccs_otl %>%
            dplyr::filter(ctry_name %in% c('New Zealand')) %>%
            dplyr::select(civ, ses, opd_b, id_j) %>%
            na.omit()

# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
civ ~ c*ses   # y <- x
civ ~ b*opd_b # y <- m

opd_b ~ a*ses # m <- x

# effects of interest
xm  := a
my  := b
xy  := c
tot := c + a*b
ind := a*b
dir := c
pme := ind/tot
'

# -----------------------------------------------
# fit model
# -----------------------------------------------

m03 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_med,
       bootstrap = 5000,
       se = 'bootstrap')

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m03, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m03,
boot.ci.type = "bca.simple") %>% 
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```

# Multilevel

## Multilevel latent covariate with SES

- 1-1-1 model (unconflated MLM)
- Preacher, K. J., Zyphur, M. J., & Zhang, Z. (2010). A general multilevel SEM framework for assessing multilevel mediation. Psychological Methods, 15(3), 209–233. https://doi.org/10.1037/a0020141
- [download link](https://www.dropbox.com/s/gxa0hj0o61fu7tg/Preacher%2C%20Zyphur%2C%20Zhang_2010.pdf?dl=1)

```{r, echo=TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_mlm <- iccs_otl %>%
            dplyr::filter(ctry_name %in% c('New Zealand')) %>%
            dplyr::select(civ, ses, opd, id_j) %>%
            na.omit()



# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
level: 1
civ ~ c1*ses # y <- x
civ ~ b1*opd # y <- m
opd ~ a1*ses # m <- x

opd ~ 0*1 

level: 2
civ ~ c2*ses # y <- x
civ ~ b2*opd # y <- m
opd ~ a2*ses # m <- x

## L1 effects
xm1  := a1
my1  := b1
xy1  := c1
tot1 := c1 + a1*b1
ind1 := a1*b1
dir1 := c1
pme1 := ind1/tot1

## L2 effects
xm2  := a2
my2  := b2
xy2  := c2
tot2 := c2 + a2*b2
ind2 := a2*b2
dir2 := c2
pme2 := ind2/tot2

# decomposition
ses_c  := tot2 - tot1
ses_w  := tot1
otl_b  := a2*b2
ses_p  := tot2 - tot1 - a2*b2
total  := ses_p + ses_w + otl_b
pme_c  := ind2/ses_c
pme_t  := ind2/total

'


# -----------------------------------------------
# fit model
# -----------------------------------------------

m04 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_mlm,
       cluster = 'id_j')

# -----------------------------------------------
# display summary
# -----------------------------------------------

lavaan::summary(m04, 
  fit.measures=TRUE,
  standardized=TRUE,
  rsquare=TRUE)

#------------------------------------------------
# indirect effects
#------------------------------------------------

lavaan::parameterEstimates(m04)%>% 
dplyr::filter(op == ':=') %>%
knitr::kable(., digits = 2)


```


# MLM estimates from all countries

## MSEM as a function

```{r, echo = TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

get_otl <- function(data){

data_model <- data

get_otl_mlm <- function(data){

# -----------------------------------------------
# prep data
# -----------------------------------------------

data_mlm <- data %>%
            dplyr::select(civ, ses, opd, id_j) %>%
            group_by(id_j) %>%
            mutate(n_j = n()) %>%
            ungroup() %>%
            # dplyr::filter(n_j > 9) %>%
            na.omit()

# -----------------------------------------------
# define model
# -----------------------------------------------

lavaan_model <-'
level: 1
civ ~ c1 * ses   # y <- x
civ ~ b1 * opd   # y <- m
opd ~ a1 * ses   # m <- x


level: 2
civ ~ c2 * ses   # y <- x
civ ~ b2 * opd   # y <- m
opd ~ a2 * ses   # m <- x


# L1 effects
xm1  := a1
my1  := b1
xy1  := c1
tot1 := c1 + a1*b1
ind1 := a1*b1
dir1 := c1
pme1 := ind1/tot1


# L2 effects
xm2  := a2
my2  := b2
xy2  := c2
tot2 := c2 + a2*b2
ind2 := a2*b2
dir2 := c2
pme2 := ind2/tot2

# decomposition
ses_c  := tot2 - tot1
ses_w  := tot1
otl_b  := a2*b2
ses_p  := tot2 - tot1 - a2*b2
total  := ses_p + ses_w + otl_b
pme_c  := ind2/ses_c
pme_t  := ind2/total
'


# -----------------------------------------------
# fit model
# -----------------------------------------------

m04 <- lavaan::sem(lavaan_model, 
       mimic = 'MPLUS',
       data = data_mlm,
       cluster = 'id_j')

#------------------------------------------------
# indirect effects
#------------------------------------------------

est <- lavaan::parameterEstimates(m04)%>% 
dplyr::filter(op == ':=')

#------------------------------------------------
# return
#------------------------------------------------

return(est)

}

#------------------------------------------------
# run through all
#------------------------------------------------

otl_est <- data_model %>%
           split(.$ctry_name) %>%
           purrr::map( ~ get_otl_mlm(data = .))  %>%
           dplyr::bind_rows(., .id = 'ctry_name')

#------------------------------------------------
# return
#------------------------------------------------

return(otl_est)

}

# -----------------------------------------------
# create estimates
# -----------------------------------------------

otl_est <- get_otl(iccs_otl)

# -----------------------------------------------
# display estimates
# -----------------------------------------------

otl_est %>%
knitr::kable(., digits = 2)

# -----------------------------------------------
# display estimates
# -----------------------------------------------

otl_est %>%
dplyr::filter(lhs %in% c('my2')) %>%
arrange(desc(est)) %>%
knitr::kable(., digits = 2)


otl_est %>%
dplyr::filter(lhs %in% c('ind2')) %>%
arrange(desc(est)) %>%
knitr::kable(., digits = 2)


otl_est %>%
dplyr::filter(lhs %in% c('pme_c')) %>%
arrange(desc(est)) %>%
knitr::kable(., digits = 2)

otl_est %>%
dplyr::filter(lhs %in% c('pme_t')) %>%
arrange(desc(est)) %>%
knitr::kable(., digits = 2)




```

# Plots

## Plot of total SES gap

```{r, echo = TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

library(ggplot2)

# -----------------------------------------------
# display estimates
# -----------------------------------------------

order_plot <- otl_est %>%
              dplyr::filter(lhs %in% c('total')) %>%
              arrange(est) %>%
              mutate(order = seq(1:nrow(.))) %>%
              dplyr::select(ctry_name, order) %>%
              dplyr::glimpse()

data_plot <- otl_est %>%
             dplyr::filter(lhs %in% c('total')) %>%
             mutate(type = case_when(
              lhs == 'total' ~ 'ses b'
              )) %>%
              mutate(alpha_w = case_when(
              pvalue >  .05 ~ .2,
              pvalue <= .05 ~ 1
              )) %>%
             dplyr::left_join(., order_plot, by = 'ctry_name') %>%
             arrange(order) %>%
             mutate(type_fct = forcats::as_factor(type)) %>%
             dplyr::glimpse()

alpha_weight <- data_plot$alpha_w


# -----------------------------------------------
# colour palette
# -----------------------------------------------

# primary
unesco_blue <- '#0077d4'
sdg4_red    <- '#c5192d'

# secondary
light_green <- '#95c11f'
purple      <- '#9f358b'
orange      <- '#f39200'

# tertiary
yellow      <- '#ffd500'
magenta     <- '#ec619f'
cyan        <- '#00b1eb'
dark_blue   <- '#004d6d'


# -----------------------------------------------
# define selected colors
# -----------------------------------------------

color_0    <- dark_blue
color_1    <- 'grey20'      # darken colour
color_2    <- sdg4_red      # light colour
plot_color <- unesco_blue

# Stacked
ses_plot <- ggplot(data_plot, 
    aes(
      colour = type_fct, 
      fill   = type_fct, 
      y      = est, 
      x      = reorder(ctry_name, order
        ))) + 
    geom_bar(
      position="stack", 
      stat="identity",
      alpha = alpha_weight
      ) +
ylim(-.5,2) +
    coord_flip() +
scale_fill_manual(values = c(unesco_blue)) +
scale_color_manual(values = c('grey20')) +
ylab('SES gap on civic knoweledge decomposition
  \nSES total = ses_w + ses_c
  \n(y, m = 0, sd = 1)') +
xlab('Countries') +
theme(
axis.text.y = element_text(colour = color_1, size = 9),
axis.text.x = element_text(colour = color_1, size = 8),
panel.background = element_blank(),
panel.grid.minor = element_blank(),
panel.grid.major = element_blank(),
panel.grid.major.y = element_line(
      size = 0.1, 
      linetype = 'solid', 
      colour = color_1
      ),          
panel.grid.major.x = element_line(
      size = 0.1, 
      linetype = 'solid', 
      colour = color_2
      ),          
axis.ticks = element_blank(),
legend.title = element_blank(),
legend.background=element_blank(),
legend.key =element_rect(fill = NA),
panel.border = element_blank(),
    axis.line.x = 
    element_line(
      size = 0.4, 
      linetype = 'solid', 
      colour = color_1),
    axis.line.y = 
    element_line(
      size = 0.4,
      linetype = 'solid',
      colour = color_1),
text = element_text(size = 10,  family = 'Avenir'),
plot.title = element_text(
  family = "Avenir", 
  size = 10, 
  hjust = 0.5,
  colour = color_1) 
    )

# -----------------------------------------------
# display plot
# -----------------------------------------------

ses_plot

```

## Plot of SES contextual

```{r, echo = TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

library(ggplot2)

# -----------------------------------------------
# display estimates
# -----------------------------------------------

order_plot <- otl_est %>%
              dplyr::filter(lhs %in% c('total')) %>%
              arrange(est) %>%
              mutate(order = seq(1:nrow(.))) %>%
              dplyr::select(ctry_name, order) %>%
              dplyr::glimpse()

data_plot <- otl_est %>%
             dplyr::filter(lhs %in% c('ses_c','ses_w')) %>%
             mutate(type = case_when(
              lhs == 'ses_c' ~ 'ses contextual',
              lhs == 'ses_w' ~ 'ses within'
              )) %>%
             mutate(stacked_order = case_when(
              lhs == 'ses_w' ~ 2,
              lhs == 'ses_c' ~ 1
              )) %>%
              mutate(alpha_w = case_when(
              pvalue >  .05 ~ .2,
              pvalue <= .05 ~ 1
              )) %>%
             dplyr::left_join(., order_plot, by = 'ctry_name') %>%
             arrange(order, stacked_order) %>%
             mutate(type_fct = forcats::as_factor(type)) %>%
             dplyr::glimpse()

alpha_weight <- data_plot$alpha_w


alpha_weight <- data_plot$alpha_w


# -----------------------------------------------
# colour palette
# -----------------------------------------------

# primary
unesco_blue <- '#0077d4'
sdg4_red    <- '#c5192d'

# secondary
light_green <- '#95c11f'
purple      <- '#9f358b'
orange      <- '#f39200'

# tertiary
yellow      <- '#ffd500'
magenta     <- '#ec619f'
cyan        <- '#00b1eb'
dark_blue   <- '#004d6d'


# -----------------------------------------------
# define selected colors
# -----------------------------------------------

color_0    <- dark_blue
color_1    <- 'grey20'      # darken colour
color_2    <- sdg4_red      # light colour
plot_color <- unesco_blue

# Stacked
ce_plot <- ggplot(data_plot, 
    aes(
      colour = type_fct, 
      fill   = type_fct, 
      y      = est, 
      x      = reorder(ctry_name, order
        ))) + 
    geom_bar(
      position="stack", 
      stat="identity",
      alpha = alpha_weight
      ) +
ylim(-.5,2) +
    coord_flip() +
scale_fill_manual(values = c(unesco_blue, orange)) +
scale_color_manual(values = c('grey20', 'grey20')) +
ylab('SES gap on civic knoweledge decomposition
  \nSES total = ses_w + ses_c
  \n(y, m = 0, sd = 1)') +
xlab('Countries') +
theme(
axis.text.y = element_text(colour = color_1, size = 9),
axis.text.x = element_text(colour = color_1, size = 8),
panel.background = element_blank(),
panel.grid.minor = element_blank(),
panel.grid.major = element_blank(),
panel.grid.major.y = element_line(
      size = 0.1, 
      linetype = 'solid', 
      colour = color_1
      ),          
panel.grid.major.x = element_line(
      size = 0.1, 
      linetype = 'solid', 
      colour = color_2
      ),          
axis.ticks = element_blank(),
legend.title = element_blank(),
legend.background=element_blank(),
legend.key =element_rect(fill = NA),
panel.border = element_blank(),
    axis.line.x = 
    element_line(
      size = 0.4, 
      linetype = 'solid', 
      colour = color_1),
    axis.line.y = 
    element_line(
      size = 0.4,
      linetype = 'solid',
      colour = color_1),
text = element_text(size = 10,  family = 'Avenir'),
plot.title = element_text(
  family = "Avenir", 
  size = 10, 
  hjust = 0.5,
  colour = color_1) 
    )

# -----------------------------------------------
# display plot
# -----------------------------------------------

ce_plot

```

## Plot of decomposition

```{r, echo = TRUE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------

library(ggplot2)

# -----------------------------------------------
# display estimates
# -----------------------------------------------

order_plot <- otl_est %>%
              dplyr::filter(lhs %in% c('total')) %>%
              arrange(est) %>%
              mutate(order = seq(1:nrow(.))) %>%
              dplyr::select(ctry_name, order) %>%
              dplyr::glimpse()

data_plot <- otl_est %>%
             dplyr::filter(lhs %in% c('ses_p','ses_w','otl_b')) %>%
             mutate(type = case_when(
              lhs == 'ses_p' ~ 'ses direct(c)',
              lhs == 'ses_w' ~ 'ses direct(w)',
              lhs == 'otl_b' ~ 'ses indirect (otl)'
              )) %>%
             mutate(stacked_order = case_when(
              lhs == 'ses_p' ~ 3,
              lhs == 'ses_c' ~ 2,              
              lhs == 'otl_b' ~ 1
              )) %>%
              mutate(alpha_w = case_when(
              pvalue >  .05 ~ .2,
              pvalue <= .05 ~ 1
              )) %>%
             dplyr::left_join(., order_plot, by = 'ctry_name') %>%
             arrange(order, stacked_order) %>%
             mutate(type_fct = forcats::as_factor(type)) %>%
             dplyr::glimpse()

alpha_weight <- data_plot$alpha_w


# -----------------------------------------------
# colour palette
# -----------------------------------------------

# primary
unesco_blue <- '#0077d4'
sdg4_red    <- '#c5192d'

# secondary
light_green <- '#95c11f'
purple      <- '#9f358b'
orange      <- '#f39200'

# tertiary
yellow      <- '#ffd500'
magenta     <- '#ec619f'
cyan        <- '#00b1eb'
dark_blue   <- '#004d6d'


# -----------------------------------------------
# define selected colors
# -----------------------------------------------

color_0    <- dark_blue
color_1    <- 'grey20'      # darken colour
color_2    <- sdg4_red      # light colour
plot_color <- unesco_blue

# Stacked
gap_plot <- ggplot(data_plot, 
    aes(
      colour = type_fct, 
      fill   = type_fct, 
      y      = est, 
      x      = reorder(ctry_name, order
        ))) + 
    geom_bar(
      position="stack", 
      stat="identity",
      alpha = alpha_weight
      ) +
ylim(-.5,2) +
    coord_flip() +
scale_fill_manual(values = c(sdg4_red, unesco_blue, orange)) +
scale_color_manual(values = c('grey20', 'grey20', 'grey20')) +
ylab('SES gap on civic knoweledge decomposition
  \nwithin, contextual, and indirect
  \n(y, m = 0, sd = 1)') +
xlab('Countries') +
theme(
axis.text.y = element_text(colour = color_1, size = 9),
axis.text.x = element_text(colour = color_1, size = 8),
panel.background = element_blank(),
panel.grid.minor = element_blank(),
panel.grid.major = element_blank(),
panel.grid.major.y = element_line(
      size = 0.1, 
      linetype = 'solid', 
      colour = color_1
      ),          
panel.grid.major.x = element_line(
      size = 0.1, 
      linetype = 'solid', 
      colour = color_2
      ),          
axis.ticks = element_blank(),
legend.title = element_blank(),
legend.background=element_blank(),
legend.key =element_rect(fill = NA),
panel.border = element_blank(),
    axis.line.x = 
    element_line(
      size = 0.4, 
      linetype = 'solid', 
      colour = color_1),
    axis.line.y = 
    element_line(
      size = 0.4,
      linetype = 'solid',
      colour = color_1),
text = element_text(size = 10,  family = 'Avenir'),
plot.title = element_text(
  family = "Avenir", 
  size = 10, 
  hjust = 0.5,
  colour = color_1) 
    )

# -----------------------------------------------
# display plot
# -----------------------------------------------

gap_plot

```


## Total and Contextual

```{r echo = TRUE, fig.width = 10, fig.asp = 1.15, out.height = '100%', fig.retina = 4, warnings = FALSE, message = FALSE, comment = FALSE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------


# -----------------------------------------------
# grid plot
# -----------------------------------------------

library(patchwork)

layout <- "
AB
AB
AB
"

p1 <- ses_plot
p2 <- ce_plot

all_gap <- p1 + p2 +
plot_layout(design = layout) +
plot_annotation(
  theme = theme(
  plot.background = element_rect(fill = "transparent", color = NA),
  panel.background = element_rect(fill = "transparent")
    )
  )

all_gap

```

## Full decomposition

```{r echo = TRUE, fig.width = 10, fig.asp = 1.15, out.height = '100%', fig.retina = 4, warnings = FALSE, message = FALSE, comment = FALSE}

# -------------------------------------------------------------------
# lavaan
# -------------------------------------------------------------------


# -----------------------------------------------
# grid plot
# -----------------------------------------------

library(patchwork)

layout <- "
AB
AB
AB
"

p2 <- ce_plot
p3 <- gap_plot


all_gap <- p2 + p3 +
plot_layout(design = layout) +
plot_annotation(
  theme = theme(
  plot.background = element_rect(fill = "transparent", color = NA),
  panel.background = element_rect(fill = "transparent")
    )
  )

all_gap

```
